Building socket-aware BPF programs

### 2018-11-01 paper/paper.tex
paper: General flow tidy up

    Using this new reference tracking ability in the verifier, we add socket
    lookup and release function calls to the BPF API, allowing BPF programs to
    safely find a socket and build logic upon the presence or attributes of a
    socket. This can be used to load-balance traffic based on the presence of a
    listening application, or to implement stateful firewalling primitives to
    understand whether traffic for this connection has been seen before. With
    this new functionality, BPF programs can integrate more closely with the
    networking stack's understanding of the traffic transiting the kernel.

@@

To allow BPF programs to retrieve and access a socket pointer, the verifier
must be made aware of when a register contains a pointer of this type, and also
how to validate pointer access. The existing verifier logic handles context
pointers in a generic manner via \verb+bpf_verifier_ops+. However, the packet
event hook points targeted in this paper already contain a context pointer of a
type that is different from \verb+bpf_sock+, so some additional logic was
required in the verifier to understand this pointer type. The patch series
associated with this paper introduced a new pointer type specific to sockets,
and linked its validation directly to the aforementioned socket verification
functions without using the verifier operations abstraction. Write operations
into the socket are rejected in the current implementation.

@@

The verifier is tasked with ensuring that all BPF program execution flow is
safe to the extents where it interacts with the BPF API. In general, BPF
programs, once verified and triggered by event hooks, will execute from the
first instruction through the final instruction for a particular path through
the program. Therefore, after execution of a BPF program containing resource
acquire and release functions which are verified by the logic above, references
to those resources should not leak.

One exception to this is when classic BPF \verb+LD_ABS+ or \verb+LD_IND+
instructions are used in the BPF program. These instructions were previously
used to provide direct packet access, with the semantics that if the access
offset exceeds the length of the packet, the BPF program would be terminated
prior to the final state. Without mitigation, this could lead to leaking of
references at runtime, even if the instructions appear to balance acquisition
and release at verification time. Newer BPF APIs provide better alternatives to
these instructions for packet access, so we disallow the use of these
instructions while holding a socket reference.

@@

Another aspect of extensibility is the ability to lookup sockets which are not
TCP or UDP. The initial RFC of this patch series proposed a single lookup
function which would choose the Layer~4 protocol based upon a field in the
tuple~\cite{sk-lookup-rfc}, however this would make it more difficult for BPF
program writers to detect the support for different protocol types at compile
time. When a socket type is unsupported, the implementation would return
\verb+NULL+, implying that there is no such socket. There is not a significant
number of supported Layer~4 socket types in the Linux stack today, and the
number is not expected to increase drastically, so it was considered simpler
and easier to split out each Layer~4 socket lookup function into an independent
helper call, so that the result could simply return a socket or \verb+NULL+ and
not need to encode a representation of other error conditions.

@@

The API retains explicit reference tracking to ensure consistency of the API
and to allow multiple underlying implementations to handle reference tracking
in a way that is safe for the implementation. For instance, for  TCP sockets
that are not governed by RCU, the networking stack uses reference counting to
manage socket memory instead, so the socket lookup and release functions would
take and release references on the socket, respectively.

@@

In cases where Linux is configured as a firewall between applications that
exist in distinct Linux kernel copies running in the network, this makes sense;
The information that is needed to perform stateful firewalling is not present
on the Linux instance that needs this information, so the information must
either be shared from the peers (assuming that the firewall orchestrator has
control over the peer endpoints), or this information must be synthesised based
upon the packets that are seen on this intermediate Linux instance. However, in
many cases this model is used when filtering traffic where one endpoint is
co-located with the Linux stack that performs the firewalling, which already
contains information about the local end of the connection. In this use case,
the connection tracking table effectively duplicates knowledge that Linux
already creates and maintains.

@@

\subsection{Filtering management traffic in forwarding devices}

@@

Using the upcoming XDP support for socket lookup, early in the driver level of
the Linux stack the traffic could be compared with the socket table and only
passed up the stack if the traffic corresponds to a local application. As this
functionality uses the socket table directly, it is always kept in sync with
the current applications without additional orchestration of BPF maps. Traffic
which is not intended to be handled locally can then be either handled by
additional BPF logic programmed in the XDP hook.

@@

This paper describes the contribution of new BPF verifier functionality in the
Linux kernel to track references to kernel resources and uses these to provide
access to socket introspection capabilities from packet handling hooks. The
introduction of reference tracking logic into the verifier provides
useful base infrastructure for supporting acquire and release semantics for
kernel resources which may prove useful for other helpers in future, and the
socket lookup API makes the packet hooks more powerful to support use cases
such as stateful firewalling or load-balancing based on the sockets that are
open on a system.

>>>>

    Using this new reference tracking ability in the verifier, we add socket
    lookup and release function calls to the BPF API, allowing BPF programs to
    safely find a socket and build logic upon the presence or attributes of a
    socket. This can be used to conditionally redirect traffic based on the
    presence of a listening application, or to implement stateful firewalling
    primitives to understand whether traffic for this connection has been seen
    before. With this new functionality, BPF programs can integrate more
    closely with the networking stack's understanding of the traffic transiting
    the kernel.

@@

To allow BPF programs to retrieve and access a socket pointer, the verifier
must be made aware of when a register contains a pointer of this type, and also
how to validate pointer access. The existing verifier logic handles context
pointers in a generic manner via \verb+bpf_verifier_ops+. The core of the
verifier is not aware of specifice pointer types that are used as the context
for the program, only how to verify the pointer access for these types.
However, the packet event hook points targeted in this paper already contain a
context pointer of a type that is different from \verb+bpf_sock+, so some
additional logic was required in the verifier to understand this pointer type.
The patch series associated with this paper introduced a new pointer type
specific to sockets, and linked its validation directly to the aforementioned
socket verification functions without using the verifier operations
abstraction. Write operations into the socket are rejected in the current
implementation. If variations in access and access type into the socket is
desired in future based upon the hook point, then it may make sense to
introduce a similar abstraction for the socket type.

@@

The verifier is tasked with ensuring that all BPF program execution flow is
safe to the extents where it interacts with the BPF API. In general BPF
programs---once verified and triggered by event hooks---will execute from the
first instruction through the final instruction for a particular path through
the program. Therefore, references to those resources should not leak if the
program has been verified.

One exception to this is when classic BPF \verb+LD_ABS+ or \verb+LD_IND+
instructions are used in the BPF program. These instructions were previously
used to provide direct packet access, with the semantics that if the access
offset exceeds the length of the packet, the BPF program would be terminated
prior to reaching the final state. Without mitigation, this could lead to
leaking of references at runtime, even if the instructions appear to balance
acquisition and release at verification time. Newer BPF APIs provide better
alternatives to these instructions for packet access, so we disallow the use of
these instructions while holding a socket reference.

@@

Another aspect of extensibility is the ability to lookup sockets which are not
TCP or UDP. The initial RFC of this patch series proposed a single lookup
function which would choose the Layer~4 protocol based upon a field in the
tuple~\cite{sk-lookup-rfc}, however this would make it more difficult for BPF
program writers to detect the support for different protocol types at compile
time. When a socket type is unsupported, the initial implementation would
return \verb+NULL+ at runtime, implying that there is no such socket. There is
not a significant number of supported Layer~4 socket types in the Linux stack
today, and the number is not expected to increase drastically, so it was
considered simpler and easier to split out each Layer~4 socket lookup function
into an independent helper call, so that the result could simply return a
socket or \verb+NULL+ and not need to encode a representation of other error
conditions. This simplifies not only the API, but also the verifier
implementation that must understand the return type of the helper functions.

@@

The API retains explicit reference tracking to ensure consistency of the API
and to allow multiple underlying implementations to handle reference tracking
in a way that is safe for the implementation. For instance, for  TCP sockets
that are not governed by RCU, the networking stack uses reference counting to
manage socket memory instead, so the socket lookup and release function
implementations must take and release references on the socket, respectively.

@@

In cases where Linux is configured as a firewall between applications that
exist in distinct Linux kernel copies running in the network, this is
intuitive. The information that is needed to perform stateful firewalling is
not present on the Linux instance that needs this information, so to understand
the state of the peers the information must be ether shared from the peers
(assuming that the firewall orchestrator has control over the peer endpoints),
or this information must be synthesised based upon the packets that are seen on
this intermediate Linux instance. However, in many cases this model is used
when filtering traffic where one peer is co-located with the Linux stack that
performs the firewalling, which already contains information about the local
end of the connection. In this use case, the connection tracking table
effectively duplicates knowledge that Linux already creates and maintains.

@@

\subsection{Conditional forwarding of management traffic in forwarding devices}

@@

Using the upcoming XDP support for socket lookup, early in the driver level of
the Linux stack the traffic could be compared with the socket table and only
passed up the stack if the traffic corresponds to a local application. As this
functionality uses the socket table directly, it is always kept in sync with
the current applications without additional orchestration of BPF maps. Traffic
which is not intended to be handled locally can then be handled by additional
BPF logic programmed in the XDP hook.

@@

This paper describes the contribution of new BPF verifier functionality in the
Linux kernel to track references to kernel resources and uses these to provide
access to socket introspection capabilities from packet handling hooks. The
introduction of reference tracking logic into the verifier provides
useful base infrastructure for supporting acquire and release semantics for
kernel resources which may prove useful for other helpers in future, and the
socket lookup API makes the packet hooks more powerful to support use cases
such as stateful firewalling or conditional redirect based on the sockets that
are open on a system.
